|

### 2021-12-16 04-Chp-Systematic-Review-Results.Rmd
End of day:
Range of updates

Using the methods outlined in the previous chapter, here I present  the results of a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides) and dementia outcomes. In addition, it examined the relationship between treatments that change blood lipid levels, such as statins, and dementia outcomes.

@@

<!-- TODO Complete! -->

@@

<!-- TODO Compare findings for two papers from Rantanen - not sure why? -->

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. To ensure the accuracy of the study design filters, a random sample of 500 records removed by the filters were screened, and no eligible records were identified. As expected, many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

@@

The PRISMA flow diagram presented in Figure \@ref(fig:prisma-flow-fig) illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described previously (presented in grey on the right of the diagram).

Common reasons for exclusion at the full-text stage included studies reporting on ineligible exposures (Number of studies (N) = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (N = `r n_wrong_outcome`; e.g. change in cognitive scores), or the use of an ineligible study design (N = `r n_study_design`).

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of Gwet's agreement coefficinet, $AC1$, were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

<!----------------------------------------------------------------------------->
(ref:agreeInter-caption) __Inter-rater agreement__ - For a subset of records, the decisions between two independent were compared. Gewt's $AC1$ was calculated from this data to assess inter-rater reliability, indicating high accuracy ($AC1$ = `r agreeInter_coeff[1]`).

<!-- TODO Include AC1 result in caption above -->

@@

@@

@@

The majority of included studies described non-randomised analyses, with the only two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on incidence of all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS), leading to complications in the synthesis of these estimates due to the potential for double-counting (see Section \@ref(sys-rev-res-AD)).

Of the `r n_nrsi_included` non-randomised studies examining treatments that modify lipid levels, all examined statin use (`r n_statins`), while a small number also reported on other non-statins agents such as fibrates (`r n_fibrates`). In the `r n_nrse_included` non-randomised studies of exposure, hypercholesterolemia (`r n_hyperchol`) and total cholesterol levels (`r n_TC`) were the most frequently reported risk factors.

@@

<!-- Many studies using electronic health records as their data source did not accurately report the diagnostic codes used to identify cases. -->

<!-- Midlife cholesterol levels were of particular interest, with several studies examining this age-range.  -->

<!--- TODO Talk a little bit more about age here --->

<!-- TODO Also need to extract the codes used in each study -->
<!-- TODO Need to make sure that the MR studies actually don't report the  -->

@@

\blandscape{}
<!----------------------------------------------------------------------------->
<!-- To create the references below, go to the helper script and copy the output of  -->

@@

<!-- TODO Sort out the fact that statins are included as an exposure in NRSE -->

<!-- TODO Ensure all abbreviations at the end of this table are in the abbreviation section at the start. -->

@@

(ref:cohortLocations-cap) __Geographical distribution of study cohorts__ - The country or countries of origin for the studies include in the review are shown below. Note that the majority of included studies were performed in the Western world, predominantly in Europe and North America.

@@

@@

<!-- TODO Deal with awkward hypercholesterolemia definition -->

@@

Finally, there were no identified Mendelian randomisation analysis examining the effect of lower lipid levels, as determined by any genetic instrument, on all-cause dementia risk.

@@

The two randomised controlled trials provided weak evidence (`r rct_statin_acd`) of an effect on statin use on all-cause dementia risk (Figure \@ref(fig:rctStatinDementiaFig)).

@@

In contrast, a meta-analysis of `r obsStatins$Dementia$n` prospective observational studies provided some evidence of a protective effect of statins use on all-cause dementia risk (`r obsStatins$Dementia$estimate`, Figure \@ref(fig:obsStatinDementiaFig)).

@@

Finally, a single Mendelian randomisation analysis was identified examining the effect of  lowered LDL-c levels on the risk of all-cause dementia via genetic inhibition of the 3-hydroxy-3-methylglutaryl-coenzyme A reductase (HMGCR), emulating statin treatment (see Section \@ref(intro-statins) for more details of the statin mechanism of action). This analysis provided weak evidence of an effect (`r mrStatin$Dementia$estimate`).

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies took a two-sample approach, making use of summary statistics from the Global Lipid Genetics Consortium (GLGC) and International Genomics of Alzheimer's Project (IGAP) consortia. Synthesis of these results would would result in a falsely precise estimate caused by multiple counting of the same participants,and as a result, no meta-analysis of these studies was performed. However, across each analysis using varying number of SNPS, no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

There were no randomised trials of the use of statins or any other lipid regulating agents  on Alzheimer's disease, though several observational studies reported on this outcome and provided evidence for a protective effect (`r n_effect(obsStatins$AD)`; Figure \@ref(fig:obsStatinADFig)).

@@

`r num_to_text(obsHyperchol$VaD$n,T)` studies reported on the association of hypercholesterolemia with vascular dementia and provided weak evidence for an effect (`r obsHyperchol$VaD$estimate`, Figure \@ref(fig:obsHyperVaD))

@@

Few studies investigated the effect of individual lipid fractions. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions,[@yoshitake1995] and similarly found minimal evidence of an effect LDL-c (`r n_effect(obsLipids$VaD_LDL,F)`), HDL-c (`r n_effect(obsLipids$VaD_HDL,F)`) or triglycerides (`r n_effect(obsLipids$VaD_TG,F)`).

@@

As noted in Section \@ref(sys-rev-characteristics) above, there was substantially less literature available on the association of my risk factors of interest and vascular dementia. There were no available randomised trials for this outcome. `r num_to_text(obsStatins$VaD$n,T)` prospective cohort studies examined statin use and vascular dementia, though meta-analysis of these studies provided weak evidence for an effect (`r n_effect(obsStatins$VaD)`; Figure \@ref(fig:obsStatinVaDFig)).

@@

As shown in the forest plots in the above sections, the risk-of-bias assessments are presented alongside their corresponding numerical result. This section presents only a brief summary of the biases observed in each study design, as s detailed discussion of the sources and directions of bias in each result is presented in Chapter \@ref(tri-heading).

@@

The risk of bias due to missing evidence in each synthesis is shown beside the overall summary diamond in each forest plot presented above. For randomised controlled trials and non-randomised studies of interventions, the risk of bias due to missing evidence was assessed to be minimal. However, there was substantial evidence that results were selectively reported in studies examining the effect of lipid fractions on dementia outcomes (Figures \@ref(fig:lipidFractionsDementia), \@ref(fig:lipidFractionsAD) and \@ref(fig:lipidFractionsVaD)).

@@

There were `r n_dr_initial` studies initially considered eligible for the dose-response meta-analysis, as they appear to provide data on risk of dementia outcomes across several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not report the relevant information needed (most commonly, the cut-off measures for each category).

@@

Investigation of potential sources of heterogeneity was complicated by two factors. In the first instance, a minority of meta-analysis (4 out of 18, `r paste0(comma(4/18*100),"%")`) included more than 10 results, the recommended minimum required for meta-regression. Secondly, poor reporting of study characteristics of interest including education level and baseline cognitive ability precluded the use of several results in a meta-regression analysis.

<!-- TODO Finish this analysis -->

Age and gender were assessed as potential causes of heterogeneity in the meta-analyses of statin use and hypercholesterolemia on all cause dementia and Alzheimer's disease, but I found weak evidence for variation in the observed effect estimates due to these factors.

Similarly, assessment of small-study effects, for which publication bias could be one potential reason, was hindered by the relatively small number of results included in a given meta-analysis. However, all analyses assessed provided weak evidence of small-study effects.

<!-- TODO Reword the above section to make it clearer -->

@@

As show in the PRISMA flow diagram (Figure \@ref(fig:prisma-flow-fig)), the number of hits returned by preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of the 18 meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the result of the meta-analysis (`r mrStatin$AD$estimate`) compared with that reported by the single published study (`r estimate(0.59,0.25,1.39,"RR")`), other than a slight increase in precision.

The other two preprints identified[@andrews2019; @zhu2017] reported on the effect of LDL-c on Alzheimer's disease using data from GLCC and IGAP consortia. As illustrated in Figure \@ref(fig:mrDuplication), these consortia were previously analysed in several published reports, and so these preprints did not add to new information to the evidence base.

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia. This discussion seeks to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise in Chapter \@ref(tri-heading).

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of differing exposure windows (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

<!-- Some evidence that age has impact on observed lipid-dementia relationship - use to link  -->

Across dementia outcomes, the majority of studies were non-randomised studies of lipids or lipid regulating agents. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the follow-up made necessary by the long latent period of the condition makes trials logistically difficult and financially expensive. Similarly, Mendelian randomisation is a comparatively new study design, and studies employing it in relation to dementia outcomes only appear in the evidence base in recent years, as illustrated in Figure \@ref(fig:typeByYear). This recent increase is likely driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

A central finding of this review is the absence of studies examining vascular dementia as an outcome, most noticeable when comparing the evidence base for statins in dementia/Alzheimer's (Figures \@ref(fig:obsStatinDementiaFig) & \@ref(fig:obsStatinADFig)) with that available for vascular dementia (Figure \@ref(fig:obsStatinVaDFig)). This is particularly interesting given that lipids and statins are strongly related to the prevention of vascular disease. A potential explanation for this observation may be publication bias or the "file-drawer effect",[@rosenthal1979] though there was weak evidence of a small-study effects for this outcome (of which publication bias is one potential cause).[@sterne2011] Similarly, only one Mendelian randomisation study examined this outcome, likely due to the absence (until recently) of vascular dementia GWAS which precludes a two-sample approach.

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on cardiovascular disease risk and reported on cognitive outcomes as one of several secondary outcomes.[@shepherd2002] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though a useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria, as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not meet the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally high. The causes and expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates were known to be missing from meta-analyses due to preferential reporting of significant results, leading to high risk of bias due to missing evidence. These missing estimates were most commonly identified through differing descriptions of the same analysis in conference abstract versus the final publication.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table."[@ancelin2013]). However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided weak evidence of an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of the included studies being conducted in the Western world (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of several studies which made use of data from the Taiwan health insurance database.

@@

<!-- TODO Check none of these use a ROB tool -->

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, to my knowledge, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomised controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published. In addition this previous review extracted the results of analyses which did not account for SNPs in the _Apo_$\mathcal{E}4$ genetic region (see the following section for a discussion of the bias this introduces).

Despite these differences in time scales and methodology, the duplication of work across systematic reviews (including this one) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch, known as an umbrella review or review-of-reviews,[@aromataris2015; @smith2011] could have been employed. This study design uses other systematic reviews rather than primary studies as the unit of analysis. Furthermore, it would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart could have been applied.

@@

* the genetic variant associates with the risk factor of interest (relevance assumption);
* the variant-exposure association has no unmeasured confounders (independence assumption);
* the variants affect the outcome only through their effect on the risk factor of interest (exclusion restriction assumption).

@@

(ref:mrAssumptions-cap) __Summary of the assumptions in Mendelian randomisation analyses__ - Mendelian randomisation analyses make three key assumptions: (I) _relevance_ - the genetic variant associates with the risk factor of interest; (II) _independence_ - the variant-exposure association has no unmeasured confounders; and (III) _exclusion restriction_ - variants affect the outcome only through their effect on the risk factor of interest (i.e. there is no horizontal pleiotropy).

@@

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist.[@davies2018] Reporting quality is an important consideration for Mendelian randomisation studies, as reflected by the recent release of the STROBE MR reporting guidelines,[@skrivankova2021] but it is a separate consideration to bias, as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was designed to assess the quality of GWAS included in a meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not address the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]

>>>>

Using the methods outlined in the previous chapter, here I present the results of a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides) and dementia outcomes. In addition, I also examined the relationship between treatments that change blood lipid levels, such as statins, and dementia outcomes.

@@

@@

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. To ensure the accuracy of the study design filters, a random sample of 500 records removed by the filters were screened, and no eligible records were identified. As expected, many of those excluded by the filters were commentaries/educational articles, or described basic science studies.

@@

A PRISMA flow diagram, presented in Figure \@ref(fig:prisma-flow-fig), illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the preprint search tool described previously (presented in grey on the right of the diagram).

Common reasons for exclusion at the full-text stage included studies reporting on ineligible exposures (Number of records (N) = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (N = `r n_wrong_outcome`; e.g. change in cognitive scores), or the use of an ineligible study design (N = `r n_study_design`).

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of Gwet's agreement coefficinet, $AC1$, were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). 
&nbsp;

<!----------------------------------------------------------------------------->
(ref:agreeInter-caption) __Inter-rater agreement__ - For a subset of records, the decisions between two independent reviewers were compared. Gewt's $AC1$ was calculated from this data to assess inter-rater reliability, indicating high accuracy ($AC1$ = `r agreeInter_coeff[1]`).

@@

Similarly, for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)).

@@

In both cases, the discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance towards exclusion.[@feinstein1990]

@@

The majority of included studies described non-randomised analyses, with the only two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on incidence of all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS). This leads to complications in the synthesis of these estimates due to the potential for double-counting (see Section \@ref(sys-rev-res-AD)).

Of the `r n_nrsi_included` non-randomised studies examining lipid-regulating agents (LRA), all examined statin use while a small number also reported on other non-statin agents such as fibrates (`r n_fibrates`). In the `r n_nrse_included` non-randomised studies of exposure, hypercholesterolemia (`r n_hyperchol`) and total cholesterol levels (`r n_TC`) were the most frequently reported risk factors.

@@

@@

\blandscape{}
<!----------------------------------------------------------------------------->

@@

@@

(ref:cohortLocations-cap) __Geographical distribution of study cohorts__ - The country or countries of origin for the studies included in the review are shown. Note that the majority of included studies were performed in westernised countries, predominantly in Europe and North America.

@@

Overall, the reporting of included studies was poor, particularly for non-randomised studies of lipids and LRA. Several studies did not report baseline characteristics such as the proportion of women in the study cohort. Additionally, some omitted important details on reported characteristics (e.g. the study reported the mean but not the standard deviation of baseline age).

@@

@@

Finally, there were no identified Mendelian randomisation analyses examining the effect of lower lipid levels, as determined by any genetic instrument, on all-cause dementia risk.

@@

The two randomised controlled trials identified provided weak evidence (`r rct_statin_acd`) of an effect on statin use on all-cause dementia risk (Figure \@ref(fig:rctStatinDementiaFig)).

@@

In contrast, a meta-analysis of `r obsStatins$Dementia$n` prospective observational studies provided some evidence of a protective effect of statin use on all-cause dementia risk (`r obsStatins$Dementia$estimate`, Figure \@ref(fig:obsStatinDementiaFig)).

@@

Finally, a single Mendelian randomisation analysis was identified examining the effect of  lowered LDL-c levels on the risk of all-cause dementia via genetic inhibition of the 3-hydroxy-3-methylglutaryl-coenzyme A reductase (HMGCR), emulating statin treatment (see Section \@ref(intro-statins) for more details of the statin mechanism of action). This analysis provided weak evidence for an effect (`r mrStatin$Dementia$estimate`).

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies took a two-sample approach, making use of summary statistics from the International Genomics of Alzheimer's Project (IGAP) consortium and the Global Lipid Genetics Consortium (GLGC). Synthesis of these results would result in a falsely precise estimate caused by multiple counting of the same participants, and as a result, no meta-analysis of these studies was performed. However, across each analysis using varying number of single nucleotide polymorphisms (SNPs), no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

There were no randomised trials of the use of statins or any other lipid regulating agents  on Alzheimer's disease, though several observational studies examined this association and provided evidence for a protective effect (`r n_effect(obsStatins$AD)`; Figure \@ref(fig:obsStatinADFig)).

@@

As noted in Section \@ref(sys-rev-characteristics) above, there was substantially less literature available on the association of the risk factors of interest with vascular dementia. `r num_to_text(obsHyperchol$VaD$n,T)` studies reported on the association of hypercholesterolemia with vascular dementia and provided weak evidence for an effect (`r obsHyperchol$VaD$estimate`, Figure \@ref(fig:obsHyperVaD)).

@@

Few studies investigated the effect of individual lipid fractions on vascular dementia. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these results found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions.[@yoshitake1995] This analysis similarly found minimal evidence of an effect of LDL-c (`r n_effect(obsLipids$VaD_LDL,F)`), HDL-c (`r n_effect(obsLipids$VaD_HDL,F)`) or triglycerides (`r n_effect(obsLipids$VaD_TG,F)`) on vascular dementia.

@@

There were no available randomised trials of statin use for the vascular dementia outcome. `r num_to_text(obsStatins$VaD$n,T)` prospective cohort studies examined statin use and vascular dementia, though meta-analysis of these studies provided weak evidence for an effect (`r n_effect(obsStatins$VaD)`; Figure \@ref(fig:obsStatinVaDFig)).

@@

As shown in the "paired" forest plots in the above sections, the risk-of-bias assessments are presented alongside their corresponding numerical result. This section presents only a brief summary of the biases observed in each study design, as detailed discussion of the sources and directions of bias in each result is presented in Chapter \@ref(tri-heading).

@@

The risk of bias due to missing evidence in each synthesis is shown beside the overall summary diamond in each "paired" forest plot presented above. For randomised controlled trials and non-randomised studies of interventions, the risk of bias due to missing evidence was assessed to be minimal. However, there was substantial evidence that results were selectively reported in studies examining the effect of lipid fractions on dementia outcomes (Figures \@ref(fig:lipidFractionsDementia), \@ref(fig:lipidFractionsAD) and \@ref(fig:lipidFractionsVaD)).

In the four analyses with greater than 10 results, the minimum required for subgroup analysis, stratification by overall risk-of-bias level did not demonstrate a difference between the "Serious" and "Moderate" risk of bias groups. The results of the test for subgroup differences is shown in the bottom left of the forest plots for these four meta-analyses (Figures \@ref(fig:obsHyperDementia), \@ref(fig:obsStatinDementiaFig), \@ref(fig:obsHyperAD) and \@ref(fig:obsStatinADFig)).

@@

There were `r n_dr_initial` studies initially considered eligible for the dose-response meta-analysis, as they appeared to provide data on risk of dementia outcomes across several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not report all of the information needed for the analysis. Most commonly, the cut-off measures for each category were missing.

@@

There was evidence of heterogeneity in several of the meta-analysis performed, as indicated by a high $I_2$ statistic. $I_2$ represents the percentage of variability in the results included in a synthesis that is due to heterogeneity rather than chance.[@higgins2003] The $I_2$ value for each analysis are provided in the forest plots above.

Investigation of potential sources of heterogeneity was complicated by two factors. In the first instance, a minority of meta-analysis (4 out of 18, `r paste0(comma(4/18*100),"%")`) included 10 or more results, the recommended minimum required for meta-regression. Secondly, poor reporting of baseline characteristics of interest precluded the use of several results in a meta-regression analysis.

Only for the sex covariate were sufficient number of results available to enable meta-regression. Across the four meta-analyses examined, the percentage of females in the study cohort was not related to the observed effect estimate (p = `r metaregHypercholDem` for hypercholesterolemia/all-cause dementia, Figure \@ref(fig:obsHyperDementia); p = `r metaregStatinsDem` for statins/all-cause dementia - Figure \@ref(fig:obsStatinDementiaFig); p = `r metaregHypercholAd` for hypercholesterolemia/Alzheimer's disease - Figure \@ref(fig:obsHyperAD); p = `r metaregStatinsAd` for statins/Alzheimer's disease - Figure \@ref(fig:obsStatinADFig)).

Similarly, assessment of small-study effects, for which publication bias could be one potential reason, was limited to these four meta-analyses including 10 or more results. However, there was weak evidence for the presence of small-study effects in each analyses assessed (see Appendix \@ref(appendix-funnel-plots)).

@@

As shown in the PRISMA flow diagram (Figure \@ref(fig:prisma-flow-fig)), the number of hits returned by preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of the 18 meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrated that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the result of the meta-analysis (`r mrStatin$AD$estimate`) compared with that reported by the single published study (`r estimate(0.59,0.25,1.39,"RR")`), other than slightly increasing the precision.

The other two preprints identified[@andrews2019; @zhu2017] reported on the effect of LDL-c on Alzheimer's disease using data from GLCC and IGAP. As illustrated in Figure \@ref(fig:mrDuplication), these consortia were previously analysed in several published reports, and so these preprints did not add new information to the evidence base.

@@

This review presented a summary of the available evidence on the association of lipids and LRA with subsequent risk of dementia outcomes. This discussion seeks to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise in Chapter \@ref(tri-heading).

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy. This suggests that these findings may be a result of differing exposure windows (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

Across dementia outcomes, the majority of studies were non-randomised studies of lipids or lipid regulating agents. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the follow-up made necessary by the long latency period of the condition makes trials logistically difficult and financially expensive. Similarly, Mendelian randomisation is a comparatively new study design, and studies employing it in relation to dementia outcomes only appear in the evidence base in recent years, as illustrated in Figure \@ref(fig:typeByYear). This recent increase is likely driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

A central finding of this review is the absence of studies examining vascular dementia as an outcome, most noticeable when comparing the evidence base for statins in dementia/Alzheimer's (Figures \@ref(fig:obsStatinDementiaFig) & \@ref(fig:obsStatinADFig)) with that available for vascular dementia (Figure \@ref(fig:obsStatinVaDFig)). This is particularly interesting given that lipids and statins are strongly related to the prevention of vascular disease. A potential explanation for this observation may be publication bias or the "file-drawer effect",[@rosenthal1979] though small-study effects (of which publication bias is one potential cause)[@sterne2011] could not be investigated for this outcome due to the small number of available results. Similarly, only one Mendelian randomisation study examined this outcome, likely due to the absence of vascular dementia GWAS which precludes a two-sample approach.

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on cardiovascular disease risk and reported on cognitive measures as one of several secondary outcomes.[@shepherd2002] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up rather than a binary dementia outcome. As such, this trial did not meet the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally high. The expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates were known to be missing from meta-analyses due to preferential reporting of significant results, leading to high risk of bias due to missing evidence. These missing estimates were most commonly identified through differing descriptions of the same analysis in conference abstract versus the final publication.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table.").[@ancelin2013] However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analyses (which provided weak evidence of an effect of lipid levels on dementia outcomes) beyond increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of the included studies being conducted in the westernised countries (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of several studies which made use of data from the Taiwan National Health Insurance database.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020; @kuzma2018risk] However, to my knowledge, this review is the first to use established domain-based assessment tools (for example, the RoB 2 tool for randomised controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or use a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published. In addition, this previous review extracted the results of analyses which did not account for SNPs in the _Apo_$\mathcal{E}4$ genetic region (see the following section for a discussion of the bias this introduces).

Despite these differences in time scales and methodology, the duplication of work across systematic reviews (including this one) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch, known as an umbrella review or review-of-reviews,[@aromataris2015; @smith2011] could have been employed. This study design uses other systematic reviews rather than primary studies as the unit of analysis. Furthermore, it would have enabled more efficient identification of a comprehensive set of primary studies relevant to this clinical question. Having defined this set, the methods which set this review apart could then have been applied.

@@

1. the genetic variant associates with the risk factor of interest (relevance assumption);
2. the variant-exposure association has no unmeasured confounders (independence assumption);
3. the variants affect the outcome only through their effect on the risk factor of interest (exclusion restriction assumption).

@@

(ref:mrAssumptions-cap) __Summary of the assumptions in Mendelian randomisation analyses__ - Mendelian randomisation analyses make three key assumptions: (1) _relevance_ - the genetic variant associates with the risk factor of interest; (2) _independence_ - the variant-exposure association has no unmeasured confounders; and (3) _exclusion restriction_ - variants affect the outcome only through their effect on the risk factor of interest (i.e. there is no horizontal pleiotropy).

@@

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this includes reporting items in their quality checklist.[@davies2018] Reporting quality is an important consideration for Mendelian randomisation studies, as reflected by the recent release of the STROBE MR reporting guidelines,[@skrivankova2021] but is a separate issue to bias as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was designed to assess the quality of GWAS included in a meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not address the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]
